{\it Maxwell's Silver Hammer:}\\ On pulsed microwave acoustic resonant viral inactivation

### 2021-02-22 documents/experiment_1.tex
More edits

\cite{Osmotic2003} 30 atmospheres, and breaks at a pressure differential of 100 atm.
T4 stress

by $\sigma = pr / 2t $,
t = 7 nm, r = 60 nm, and p =

@@

\begin{autem}
As noted, this field reached a reasonable level of quality. 1951. Some standards.
By using yet another design, with a new set of unknown artifacts, we are actually hampering this field;
and so we encourage anyone who may wish to replicate this result not to use this system.
In this case, we are limited by budget - primarily the power amplifiers we have access to.
\end{autem}

>>>>

@@

\begin{autem}
As noted, this field reached a reasonable level of quality. 1951. Some standards.
By using yet another design, with a new set of unknown artifacts, we are actually hampering this field;
and so we encourage anyone who may wish to replicate this result not to use this system, but one of the better-characterized systems mentioned below.
In this case, we are limited by budget - primarily the power amplifiers we have access to.
\end{autem}

### 2021-02-22 documents/experiment_1_data.tex
More edits

\begin{center}
	\begin{tabular}{ |c|c|c| } 
		\hline
		 & col2 & col3 \\
		\hline
		\multirow{3}{4em}{Multiple row} 
		& cell2 & cell3 \\ 
		& cell5 & cell6 \\ 
		& cell8 & cell9 \\ 
		\hline
	\end{tabular}
\end{center}

>>>>

\begin{table}[h!]
	\centering
	\begin{tabular}{ |c|c|c| } 
		\hline
		Exposure & kilocounts per 10s integration & Note \# \\
		\hline
		Autoclaved phage (0.2 mL) & 30 & \\ 
		Autoclaved p. (50 uL) & 32 & Run 8\\ 
		\hline
		Phage culture 'blank' & 8 & \\ 
		Water & 1 & \\
		\hline
		\multicolumn{3}{|c|}{\ntilde 10 or 20 us  } \\
		\hline
		 & \\ 
		\hline
		\multicolumn{3}{|c|}{Sweep from 6 to 12 GHz, average approximately 150 V/m, 10 or 20 us CW pulses} \\
		\hline
		\multicolumn{3}{|c|}{\ntilde 0.05 mm, \ntilde } \\
		\hline
		\multicolumn{3}{|c|}{\ntilde 0.2 mm channel, 0.2 uL/s, various pulse counts and termination} \\
		\hline 
		 Blank    & 8 & Li test 1\\
		 +           & 8 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 +           & 7 & Li test 1\\
 		 + \& shaper & 7 & Li test 1\\
 		 \hline
		\multicolumn{3}{|c|}{\ntilde 0.05 mm channel, \ntilde 0.75 MV/m, 0.2 uL/s, various pulse counts and termination} \\
		\hline
		Sham & 5 & Smaller \\
		+ & 6 & Smaller \\ 
		+ & 6 & Smaller  \\ 
		\hline
		\multicolumn{3}{|c|}{\ntilde 0.05 mm channel, \ntilde 0.75 MV/m, 0.2 uL/s, basic buffer, but pH may not have been 10.6} \\
		\hline
		+ & 11 & Weakened \\
		+ (no term) & 11 & Weakened \\ 
		Sham & 12 & Weakened  \\ 
		\hline
	\end{tabular}
\end{table}

Data from testing on bacteriophage. Data rounded (0.5 and above up) to the nearest thousand photon counts per 10 seconds. %thoroughly characterized.

Background ('blank', in notes) readings measure the amount of DNA present in the phage culture tube before any treatment.

Repeatability between test conditions was not terrific. The source of the \ntilde 5000 count variation between some equivalent samples is not known. However, in no case did any tube approach the 22 to 30 kcounts of the autoclave.

Stock culture had titer $3\times 10^9$ T4r.

The fluor-meter printed a constant stream of results, all within the random error between capture windows $\sigma=0.5$ kcounts. Only after most of the data was taken was it realized how dumb this was.

Results aggregated over the course of several months.

In all cases, a thermal effect was implausible; the total energy in the pulse train precluded heating above the capsid melting temperature.\cite{Effects1951}

In an early iteration of the experiment, the firmware and data analysis scripts randomly generated blinded sham and autoclave assignments and automatically treated the samples accordingly. It also printed blinded templates for pipetting. This was done partly in the hopes that massively oversampling the microwave absorption spectrometer across the range of physical tolerances would highlight any resonances.\cite{first2000}

However, as spectrometry was abandoned, this non-deterministic property caused significant issues while troubleshooting the fluorescence detection system, and \st{expedience} \st{\textit{laziness}} prompted us to disable this system, violating best practices.

A thin layer of silicone conformal coating was applied to the 0.2 mm waveguide to avoid the copper ion artifact, but this could not be done on the 0.05 mm waveguide.

### 2021-02-22 documents/summary.tex
More edits

Because the effect is so remarkable, we have tried to come up with ways of explaining the observed results without invoking a non-thermal effect:
\begin{itemize}
	\item The use of microwave absorption spectroscopy techniques that have previously shown resonance-like artifacts\cite{Resonances1987}. An effective charge of $q=10^7 e^-$ was found, which is very large for a $10^9$ atom structure; genome and protein charge can hardly explain $10^5 e^-$, which would be reduced further by ionic screening. The potential energy of such a sphere of charge exceeds the metabolic burden to build a virus by a factor of 10.
	\item Infrared thermography only determines the surface temperature. Based on heating patterns seen in other experiments\cite{Effects1950}, the observed 7 C temperature rise does not seem to conclusively rule out a thermal effect. 
	\item A lack of free-space dosimetry or simulation of power density means that the frequency-selective inactivation could have been due to the frequency response of the power amplifier, antenna gain, and absorption of the cuvette.
\end{itemize}

A few well-established non-resonant non-thermal effects have already been put to practical use. Electroporation is a common laboratory procedure where an enormous electric field causes ions in solution to diffuse across a host cell's membrane, reversibly rendering it permeable (among other, more subtle mechanisms; the minutia of poration are extraordinarily complex and beyond the author's understanding\cite{Theoretical2007}). Irreversible electroporation\cite{Irreversible2013} has recently been used in the clinic\cite{Nonthermal2013} for tumor ablation.% In vitro there is evidence to suggest that sub-cellular effects take place under truly tremendous fields.

@@

In our cursory testing, Bacteriophage T4 did not appear to be lysed by microsecond pulses between 6 and 12 GHz at about 70 V/m, nor, exploring non-resonant inactivation, did nanosecond pulses up to 1 MV/m appear to have any effect. Note that neither of these regimes overlap with the results of the previous studies.

@@

@@

Electrorotation\cite{Electrorotation1988}\cite{Electrorotation1997} of viruses has been reported\cite{Analysis2004}\cite{New1999}\cite{comprehensive2001}, but as yet only in a non-destructive low-speed sub-synchronous regime for sensing, using optical means to detect rotation. Dielectrophoresis and electrostatic trapping of viruses has also been reported.

@@

Background-free photothermal optical \cite{Microwave1993a}\cite{Broadband1988} >\\ time-domain spectroscopy\cite{Time2003}\cite{Dielectric2004}\cite{Microwave1994} >\\ Raman or other non-photothermal optical (\cite{optical1983} $\neq$ [\cite{Resonances1987} \cite{Dielectric1989}]) and also \footnote{see Taylor \cite{mechanisms1981} citing unpublished, personal communication} (but we have no understanding of how such artifacts could form) >\\ VNA CW absorption or reflection (\cite{Microwave1982} $\neq$ [\cite{Resonances1987} \cite{Dielectric1989}]) \cite{Substitution1982} \cite{Millimeter1980} \footnote{Anywhere we liked};

@@

FDTD simulation of exposure geometry and sample heating is now usually possible with free, open-source software\footnote{PyTorch FDTD library from \cite{Highly2019} was used in our study with some modification, but \cite{CUDAbased2019} and \cite{openEMS} are also excellent} and should be performed. As above, metal ions from electrodes, UV exposure, and polymer surface effects\cite{Effect1994a} must be ruled out by experimental design. The only artifact-free appear to be 1. Remote exposure, actively cooled OR pulsed exposure with sufficiently short pulses to rule out heating, with.

If the effect exists, it appears to be impossible at any input power to externally drive a 10 GHz resonator in the lungs because of the layer of conductive muscle that shields the chest cavity (see supplemental). There is precedent for bronchoscopic \cite{Flexible2019}\cite{Antenna2018} or more invasive methods. Dispersion and Brillouin precursors do offer a significant improvement in loss through tissue, and the required electrical waveforms appear to be synthesizeable by variable-impedance or stub transmission-line pulse shaping\cite{Arbitrarya} . A sawtooth might be the optimal waveform (see supplemental). However, obtaining 10 GHz tones deep in the body is still very hard.

Although not sensitive to changes in infectivity, as suggested by \cite{Quantification2020}, amplification-free fluorescence dsDNA detection using photon-counting plate readers using a 1:1 mix of viral solution with 1/4000 Biotium GelGreen (a safe Acridine Orange (N-alkylacridinium) based flurophore) is a very convenient (stable over long periods of time, requires no host culture) and sufficiently sensitive (titers down to $10^8$ of T4 (170 kbp dsDNA)) assay both for capsid damage and, by melting the capsid with an autoclave, for general virus titer estimation, far faster than even high-throughput plaque assays. More useful enveloped viral surrogates like Pseudomonas phage Phi 6 are RNA and a lower sensitivity can be expected.

@@

If CCD sensors with a, ImageJ \cite{Image2012} A silicon photomultiplier may be sufficient in some cases.

However, these initial tests were marred by use of a domestic fridge to store phage, which unexpectedly froze regularly overnight, apparently reducing phage titer, along with our accidental use of T4r+, which is not an "improved" T4r, as we believed, but rather a mutation away from rapid-lysis, meaning that the lysis behavior is far slower (see \cite{Spontaneous1946}) and, in our testing, lysis of T4r+ only occurred on plaque assay plates (where culture growth and log stage infectivity is maintained for many days).

Nephelometry (right-angle light scattering measurement) and turbidimetric (transmission) using CdS and LtoF sensors; despite being visible to the naked eye "clean" 2 hour E. coli cultures do not .  Turbidimetry Impedimietric In addition, removing the necessitating 
E. coli produces the enzyme $\beta$-galactosidase to break lactose into glucose for metabolism. luciferase Promega Beta-Glo\cite{rapid2014}. However, the one part contains a proprietary detergent to lyse cells.

While undoubtedly useful for quantifying $\beta$-galactosidase, because of ambiguities in the rise and decay of the luminescence signal, short shelf life when two components are mixed, storage and reagent mixing requirements, and relatively high cost, Beta-Glo compared unfavorably to direct fluorescence for this application.

@@

In both of our final experimental setups, non-grounded, single-sided coplanar waveguides were used as exposure chambers, inspired by \cite{Microwave2007}. A programmable syringe pumped 50 uL of fluid at a constant rate through one of the coplanar channels. Simulations showed a significant decrease in field. \cite{Microchamber2011} and \cite{Characterization2012} provide a good review of pulse exposure cells and a design which is much better than ours and should be used instead. Such cells should be used rather than our crude design. \footnote{The utility wcalc was used extensively}.

@@

We are particularly excited by the prospect of using picosecond time-domain measurements not only to perform steady state dielectric measurements or monitor the input pulse, as is now common practice, but to resolve the internal effects of a single high-voltage pulse on the virus in real time. Scan-converter based transient digitizers, such as the Tektronix SCD-5000 series, appear to be the lowest-cost method of obtaining such THz sampling bandwidths.

>>>>

Because the effect is so remarkable, we have tried to come up with ways of explaining the observed results without invoking a non-thermal effect:
\begin{itemize}
	\item The use of microwave absorption spectroscopy techniques that have previously shown resonance-like artifacts\cite{Resonances1987}. An effective charge of $q=10^7 e^-$ was found, which is very large for a $10^9$ atom structure.  Genome and protein charge can hardly explain $10^5 e^-$, which would be reduced further by ionic screening. The potential energy of such a sphere of charge would exceed the metabolic burden to build a virus by a factor of 10.
	\item Infrared thermography only determines the surface temperature. Based on heating patterns seen in other experiments\cite{Effects1950}, the observed 7 C temperature rise does not seem to conclusively rule out a thermal effect. 
	\item A lack of free-space dosimetry or simulation of power density means that the frequency-selective inactivation could have been due to the frequency response of the power amplifier, antenna gain, and absorption of the cuvette.
\end{itemize}

A few well-established non-resonant non-thermal effects have already been put to practical use. Electroporation is a common laboratory procedure where an enormous electric field causes ions in solution to diffuse across a host cell's membrane, capacitively charging it, causing a change in lipid conformation, reversibly rendering it permeable (among other, more subtle mechanisms; the minutia of poration are extraordinarily complex and beyond the author's understanding\cite{Theoretical2007}). Irreversible electroporation\cite{Irreversible2013} has recently been used in the clinic\cite{Nonthermal2013} for tumor ablation.% In vitro there is evidence to suggest that sub-cellular effects take place under truly tremendous fields.

@@

Parametrizing based on all-atom simulations of small capsid segments or proteins \cite{Elucidating2009}. DNA with beads restrained by Lennard-Jones forces \cite{Communication2013}. Finite-element analysis \cite{Finite2008}; some Poisson-Boltzmann solvers

GROMACS simulations of liposomes, trajectories fourier transformed with \cite{TRAVIS2011}

In our cursory testing, Bacteriophage T4 did not appear to be lysed by 10 microsecond pulses between 6 and 12 GHz at about 70 V/m, nor, exploring non-resonant inactivation, did nanosecond pulses up to 1 MV/m appear to have any effect. Note that neither of these regimes overlap with the results of the previous studies.

@@

\cite{MICROWAVEINDUCED1975}

@@

\cite{Forced2000} Electrorotation\cite{RotatingFieldInduced1982}\cite{Electrorotation1988}\cite{Electrorotation1997} of viruses has been reported\cite{Analysis2004}\cite{New1999}\cite{comprehensive2001}, but as yet only in a non-destructive low-speed sub-synchronous regime for sensing, using optical means to detect rotation. Dielectrophoresis and electrostatic trapping of viruses has also been reported.

Rotational drag force can be found from MD simulation\cite{Evaluating2017}. A 30 nm lipid liposome using GROMACS' torque measurement feature, kindly implemented by Kutzner \cite{Keep2011}, yielded a , whereas rupture From \cite{Spinning2008}, the upper bound on the limiting angular frequency for a homegenous elastic sphere is at most $2.236 \ \bar \omega$, $$\bar \omega = \sqrt{3}\sqrt{\frac{Y}{\rho a^2}}$$ with radius a, mass density $\rho$, yield stress Y.

@@

Background-free photothermal optical \cite{Microwave1993a}\cite{Broadband1988} >\\ time-domain spectroscopy\cite{Time2003}\cite{Dielectric2004}\cite{Microwave1994} >\\ Raman or other non-photothermal optical (\cite{optical1983} $\neq$ [\cite{Resonances1987} \cite{Dielectric1989}]) and also \footnote{see Taylor \cite{mechanisms1981} citing unpublished, personal communication}, but we have no understanding of how such artifacts could form >\\ VNA CW absorption or reflection (\cite{Microwave1982} $\neq$ [\cite{Resonances1987} \cite{Dielectric1989}]) \cite{Substitution1982} \cite{Millimeter1980}.

In our experiments with our (admittedly unstable) we were able to produce many candidate resonances with perfectly justifiable modifications to post-processing. \cite{Reanalysis2020}\cite{statistical2020}. which raises interesting questions regarding software quality in research. Synthetic data

@@

FDTD simulation of exposure geometry and sample heating is now usually possible with free, open-source software and should be performed. PyTorch FDTD library from \cite{Highly2019} was used in our study with some modification, but \cite{CUDAbased2019} and \cite{openEMS} are also excellent. As above, metal ions from electrodes, UV exposure, and polymer surface effects\cite{Effect1994a} must be ruled out by experimental design. The only artifact-free appear to be 1. Remote exposure, actively cooled OR pulsed exposure with sufficiently short pulses to rule out heating, with.

If the effect exists, it appears to be impossible at any input power to externally drive a 10 GHz resonator in the lungs, partly because of the layer of conductive muscle that shields the chest cavity (see supplemental). Open-ended waveguides\cite{142018} or water bolus allow efficient coupling to the body. There is precedent for bronchoscopic \cite{Flexible2019}\cite{Antenna2018} or more invasive methods. Dispersion and Brillouin precursors do offer a significant improvement in loss through tissue, and the required electrical waveforms appear to be synthesizeable by variable-impedance or stub transmission-line pulse shaping\cite{Arbitrarya} . A sawtooth might be the optimal waveform (see supplemental). However, obtaining 10 GHz tones deep in the body is still very hard.

What quantity is referred to; induced dipole due to polarization or ionic.

T4 was selected because of large amount of history. T4 does not have a nucleocapsid; the charge distribution is likely stronger than. 
Phi6 \footnote{An interesting}

30 atmospheres. When phage is rapidly immersed in solution of high osmotic strength, the active system that maintains osmotic equilibrium is overwhelmed, and the resulting force ruptures the capsid. \cite{Osmotic2003}, and breaks at a pressure differential of 100 atm. By $\sigma = pr / 2t $ t = 5.7 nm\cite{Head1988} and r=60 nm, capsid strength should be about 53 MPa, less than the 100 to 300 MPa cited by \cite{Bacteriophage2004}.
%((100 atm * 60 nanometers) / (2*(5.7 nanometers))) to MPa

The original Yang paper used plaque assays and RT-PCR to measure envelope damage; the quantification of viral titer is a solved problem.

Although not sensitive to changes in infectivity, we used an amplification-free fluorescent dsDNA detection method, as suggested by \cite{Quantification2020}, using a custom photon-counting plate reader and a 1:1 mix of viral solution with 1/4000 Biotium GelGreen (a safe Acridine Orange (N-alkylacridinium) based flurophore).

This was a very convenient (stable over long periods of time, requires no host culture) and sufficiently sensitive (titers down to $10^8$ of T4 (170 kbp dsDNA)) assay both for capsid damage and, by melting the capsid with an autoclave, for general virus titer estimation, far faster and less laborious than high-throughput plaque assays evaluated\cite{Simple2018}\cite{Streamlining2018}. More useful enveloped viral surrogates like Phi6 are RNA and a lower sensitivity can be expected.\cite{Selected}

@@

PMT dark counts and ambient light were removed by synchronous photon counting - add when light is on, subtract when light is off.  This cannot be done with luminescence. Further filtering could be done by fluorescence lifetime discrimination.

If CCD sensors with a, ImageJ \cite{Image2012}. A silicon photomultiplier will probably be sufficient in most cases.

A number of other assays were evaluated. However, these initial tests were marred by use of a domestic fridge to store phage, which unexpectedly froze regularly overnight, apparently reducing phage titer, along with our accidental use of T4r+, which is not an "improved" T4r, as we believed, but rather a mutation away from rapid-lysis, meaning that the lysis behavior is far slower (see \cite{Spontaneous1946}). In our testing, lysis of T4r+ only occurred on plaque plates (where culture growth and log stage infectivity is maintained for many days).

Nephelometry (right-angle light scatter detection) and turbidimetry (transmission) using CdS and LtoF sensors; despite being visible to the naked eye "clean" 2 hour E. coli cultures 1.5 mL. Turbidimetry Impedimietric In addition, removing the necessitating 
E. coli produces the enzyme $\beta$-galactosidase to break lactose into glucose for metabolism. luciferase Promega Beta-Glo\cite{rapid2014}. However, one component contains a proprietary detergent designed to lyse cells and expose the galactosidase, counterproductive for phage use. Perhaps the detergent is not effective on bacteria, or the paper substrate that was used is sufficient. The detergent could perhaps be removed with BioBeads if this is desired.

While undoubtedly essential and very effective for its intended purpose (quantifying lac operon activity), because of ambiguities in the rise and decay of the luminescence signal, short shelf life when the two components are mixed, concerns regarding the detergent, and cost, Beta-Glo compared unfavorably to direct fluorescence or X-Gal for this application.

@@

In both of our final experimental setups, non-grounded, single-sided coplanar waveguides were used as exposure chambers, inspired by the CPWG heater of \cite{Microwave2007}. \cite{Nanosecond2006} describe a very similar arrangement. A programmable syringe pumped 50 uL of fluid at a constant rate through one of the coplanar channels. FDTD simulations showed a significant decrease in field. \cite{Microchamber2011} and \cite{Characterization2012} provide a good review of pulse exposure cells and a design which is much better than ours and should be used instead. \cite{Electromagnetic1989} \footnote{The utility wcalc was used extensively}.

@@

We are particularly excited by the prospect of using picosecond time-domain measurements not only to perform steady state dielectric measurements or monitor the input pulse, as is now common practice, but to resolve the internal effects of a single high-voltage pulse on the virus in real time. Scan-converter based transient digitizers, such as the Tektronix SCD-5000 series, appear to be the lowest-cost method of obtaining the >0.5 THz sampling bandwidths required, when the tube bandwidth is deconvolved.
